### 2019-03-22 paper/quarters_apj.tex
apj format

\begin{deluxetable}{cc}
\tablecaption{Targets observed conventionally for one or more quarters.}
\tablehead{\colhead{Object} & \colhead{Quarters}}
\startdata
HD~174020 & Q2, 6, 10, 14            \\
HD~175841 & Q11-12, 14-16, SC Q3     \\
HD~176582 & Q12-13                   \\
HD~178090 & Q1, 3, 10                \\
HD~180682 & Q0, 3, 7                 \\
HD~181069 & Q1, 10, 13, 14, 17       \\
HD~181878 & Q14-17                   \\
HD~182694 & Q2                       \\
HD~183124 & Even Quarters            \\
HD~185351 & Q1-3; SC Q16             \\
HD~186155 & Q1                       \\
HD~187217 & Q14-17                   \\
HD~188252 & Q13                      \\
HD~189013 & SC Q3                    \\
V380~Cyg  & Q11; SC Q7, 9, 10, 12-17 \\
V819~Cyg  & Q14, 16, 17    \\
\enddata
\tablecomments{Some smear targets were observed conventionally for one or more quarters. 
These are listed below, with SC denoting quarters that were observed in short cadence mode and all others in long cadence.  \label{quarters}}
\end{deluxetable}